Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fates of human H3N2 influenza virus lineages

### 2018-01-18 paper/manuscript.tex
moved up `\begin{document}` command to enable `latexdiff` tracking
Jesse edits to Abstract

\title{Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fate of human H3N2 influenza virus lineages}

@@

\begin{abstract}
% no more than 250 words. (170 now)
Human influenza virus rapidly accumulates mutations in its major surface protein hemagglutinin (HA).
Forecasting the evolutionary success of influenza virus strains requires understanding how these mutations affect the protein's functionality and antigenicity.
Here we experimentally measure the effects on viral growth in cell culture of all single amino-acid mutations to the HA from a recent human H3N2 influenza virus strain. 
We show that these measurements are informative for understanding viral evolution in nature, since mutations in evolutionarily successful H3N2 viral lineages are measured to generally be more favorable than mutations in lineages that die out.
We also compare our measurements to prior data for an H1 HA, and find that the amino-acid preferences of many sites have substantially shifted between H1 and H3 HAs.
For instance, there is much less of a disparity in mutational tolerance between the head and stalk domains for the H3 HA.
Overall, our work suggests that experimental measurements of mutational effects might be leveraged to help forecast the evolutionary fates of viral lineages in nature.
\end{abstract}

>>>>

\title{Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fates of human H3N2 influenza virus lineages}

@@

\begin{abstract}
% no more than 250 words. (170 now)
Human influenza virus rapidly accumulates mutations in its major surface protein hemagglutinin (HA).
The evolutionary success of influenza virus lineages depends on how these mutations affect HA's functionality and antigenicity.
Here we experimentally measure the effects on viral growth in cell culture of all single amino-acid mutations to the HA from a recent human H3N2 influenza virus strain. 
We then show that mutations in evolutionarily successful H3N2 viral lineages are measured to generally be more favorable than mutations in lineages that die out.
Therefore, despite the well-known caveats about cell-culture measurements of viral fitness, such measurements can still be informative for understanding virus evolution in nature.
We also compare our measurements for an H3 HA to similar data previously generated for a distantly related H1 HA, and find substantial differences in which amino acids are preferred at many sites.
For instance, the H3 HA has less disparity in mutational tolerance between the head and stalk domains than the H1 HA.
Overall, our work suggests that experimental measurements of mutational effects can be leveraged to help forecast the evolutionary fates of viral lineages in nature --- but only when the measurements are made on a viral strain similar to the ones being studied in nature.
\end{abstract}
